MT @HanBaldwin: Fightin OOVs in German \#twitter

### 2013-08-06 darmstadt_gscl_2013/article_twitter_normalization.tex
small updates in article

Based on these two classifications, recently proposed approaches to
the NTN task could be grouped together as follows.

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}) etc. With regard to the type of
information inference, practically all of these methods were
supervised with the exceptions of Cook and Stevenson (\cite{cook}) who
claimed to use an unsupervised technique, and Sproat et
al. (\cite{sproat}) who described both a supervised and an
unsupervised technique.

Starting from the second half of the 2000s, the raising influence and
improved quality of machine translation tools lead to the development
of NTN technologies which used broader levels of segmentation. In
\cite{aw}, Aw et al. suggested a supervised statistical system for
normalization of SMS-messages which operated on automatically aligned
phrases A few years later, Clark and Araki (\cite{clark-araki})
described a purely rule-based method which used mappings of
non-standard words and phrases to their corresponding standard
language forms.

As noted by Kobus et al. (\cite{kobus}), NTN methods relying on either
graphematic or phrasal segments usually revealed complementary
strengths and weaknesses. This notion led NLP scientists to the idea
that by incorporating multiple levels of the language into one NTN
system the total performace of the whole system would improve as
different sources of information would benefit from each other. As a
consequence of this, a wealth of combined techniques emerged in the
past few years. Among these we should especially mention works by
Kobus et al. (\cite{kobus}), Kaufmann (\cite{kaufmann}), Han and
Baldwin (\cite{han}), and Oliva et al. (\cite{oliva}). The majority of
these systems used the whole range of segmentation levels from
phonographematatic to phrasal one, and in many cases they also applied
different knowledge inference mechanisms to different levels of the
language.

@@

In Table \ref{table:mrd}, percentage figures for each of the above
subgroups are shown. We have considered OOV-distributions for both
\texttt{hunspell} and \texttt{TreeTagger}. For each of them, we
estimated the percentage of a particular subclass with regard to the
total number of occurrences of all OOV-tokens (column ``\totaloov{}'')
as well as with regard to their percentage rate in the list of only
unique OOVs disregarding the frequencies (column ``\uniqoov{}'').
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Objective limitedness of MRD''\label{table:mrd}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    regular German words & 7.82 & 8.78 & 2.8 & 3.49\\
    compounds & 1.21 & 2.42 & 2.51 & 4.55\\
    abbreviations & 3.98 & 4.77 & 3.27 & 3.44\\
    interjections & 5.95 & 4.54 & 5.58 & 4.29\\
    person names & 4.73 & 6.41 & 2.32 & 3.47\\
    geographic locations & 1.5 & 2.53 & 1.16 & 1.88\\
    company names & 2.27 & 2.84 & 4.35 & 3.01\\
    product names & 2.13 & 2.57 & 2.45 & 3.23\\
    newly coined terms & 1.35 & 1.31 & 3.33 & 2.38\\
    loanwords & 3.68 & 4.03 & 3.29 & 2.87\\
    foreign words & 11.5 & 13.76 &  9.57 & 10.91\\\hline
    {\bfseries total} & 46.12 & 53.96 & 40.63 & 43.52\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

@@

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are on the one hand either
relatively frequent omissions of characters made by users or even more
often automatic truncation of too long messages which is performed by
Twitter service itself.\footnote{As is generally known, Twitter
  imposes a strong restriction on the length of posted messages which
  can be no longer than 140 characters. Upon exceeding this length,
  tweets get automatically truncated to the maximal allowed length.}

Finally, the third group -- ``Stylistic specifics of text genre'' --
was subdivided into the subgroups:
\begin{enumerate}
  \item hashtags, e.g. \textit{\#Kleinanzeigen};
  \item @-tokens, e.g. \textit{@ZDFonline};
  \item links, e.g. \textit{http://t.co};
  \item smileys, e.g. \textit{:-P};
  \item slang, e.g. \textit{OMG}.
\end{enumerate}
according to the formal or lexical class which these tokens belonged
to. Additionally, we also considered as \textit{slang} spelling
variants of standard language words which could be regarded as their
corresponding colloquial equivalents like, for example, \textit{ned}
instead of \textit{nicht} or \textit{grad} instead of \textit{gerade}.

These cases were marked by
us both as misspellings and as slang. A detailed statistics on the
above subgroups of class 3 is shown in Table \ref{table:style}:
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Stylistic specifics of text genre''\label{table:style}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    hashtags & 7.35 & 6.18 & 13.06 & 10.59\\
    @-tokens & 13.02 & 20.23 & 16.19 & 21.91\\
    links & 2.43 & 0.4 & 4.89 & 6.07\\
    smileys & 2 & 0.73 & 6.88 & 1.2\\
    slang & 16.22 & 5.27 & 6.94 & 4.77\\\hline
    {\bfseries total} & 41.02 & 32.81 & 47.96 & 44.54\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

A striking outlier of 16.22~\% for slang in column 1 of the table is
due to the fact that the word ``RT'' which occurred 1,235 and was by
far the most frequent OOV in our data was recognized as
out-of-vocabulary by \texttt{hunspell} but not recognized as such by
\texttt{TreeTagger}.

This knowledge about composition of unknown and ill-formed words
should now help us select or devise techniques for recognizing and
correcting such tokens in cases where it is needed. These procedures
are described in the next Section of the article.

\section{Text Normalization Procedure}\label{normalization}
\subsection{Replacement of Twitter-Specific Phenomena}
%
\subsection{Restoration of Umlauts}
%
\subsection{Squeezing of Elongated Words}
%
\subsection{Translation of Slang Idioms}
%
\section{Evaluation}\label{evaluation}

@@

>>>>

According to these two classification criteria, recent approaches to
the NTN task could be grouped together as follows.

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}), Han and Baldwin (\cite{han})
etc. With regard to the type of information inference, practically all
of these methods were supervised with the exceptions of Cook and
Stevenson (\cite{cook}) who claimed to use an unsupervised technique,
and Sproat et al. (\cite{sproat}) who described both a supervised and
an unsupervised method.

Starting from the second half of the 2000s, the raising influence and
improved quality of machine translation tools led to the development
of NTN technologies which used broader levels of segmentation. In
\cite{aw}, Aw et al. suggested a supervised statistical system for
normalization of SMS-messages which operated on automatically aligned
phrases A few years later, Clark and Araki (\cite{clark-araki})
described a purely rule-based method which used mappings of
non-standard words and phrases to their corresponding standard
language forms.

As noted by Kobus et al. (\cite{kobus}), NTN methods relying on either
graphematic or phrasal segments usually revealed complementary
strengths and weaknesses. This notion led NLP scientists to the idea
that by incorporating multiple levels of the language into one NTN
system the total performace of the whole system would improve as
different sources of information would benefit from each other. As a
consequence of this, a wealth of combined techniques emerged in the
past few years. Among these we should especially mention works by
Kobus et al. (\cite{kobus}), Kaufmann (\cite{kaufmann}), and Oliva et
al. (\cite{oliva}). The majority of these systems used the whole range
of segmentation levels from phonographematatic to phrasal one, and in
many cases they also applied different knowledge inference mechanisms
to different levels of the language.

@@

In Table \ref{table:mrd}, percentage figures for each of the above
subgroups are shown. We have considered OOV-distributions for both
\texttt{hunspell} and \texttt{TreeTagger}. For each of them, we
estimated the percentage of a particular subclass with regard to the
total number of occurrences of all OOV-tokens (column ``\totaloov{}'')
as well as with regard to their percentage rate in the list of only
unique OOVs disregarding frequencies (column ``\uniqoov{}'').
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Objective limitedness of MRD''\label{table:mrd}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    regular German words & 7.82 & 8.78 & 2.8 & 3.49\\
    compounds & 1.21 & 2.42 & 2.51 & 4.55\\
    abbreviations & 3.98 & 4.77 & 3.27 & 3.44\\
    interjections & 5.95 & 4.54 & 5.58 & 4.29\\
    person names & 4.73 & 6.41 & 2.32 & 3.47\\
    geographic locations & 1.5 & 2.53 & 1.16 & 1.88\\
    company names & 2.27 & 2.84 & 4.35 & 3.01\\
    product names & 2.13 & 2.57 & 2.45 & 3.23\\
    newly coined terms & 1.35 & 1.31 & 3.33 & 2.38\\
    loanwords & 3.68 & 4.03 & 3.29 & 2.87\\
    foreign words & 11.5 & 13.76 &  9.57 & 10.91\\\hline
    {\bfseries total} & 46.12 & 53.96 & 40.63 & 43.52\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

@@

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are either relatively frequent
omissions of characters made by users or, even more often, automatic
truncations of too long messages which are performed by Twitter
service itself.\footnote{As is generally known, Twitter imposes a
  strong restriction on the length of posted messages which can be no
  longer than 140 characters. Upon exceeding this length, tweets get
  automatically truncated to the maximal allowed length.}

Finally, the third group -- ``Stylistic specifics of text genre'' --
was subdivided into the subgroups:
\begin{enumerate}
  \item @-tokens, e.g. \textit{@ZDFonline};
  \item hashtags, e.g. \textit{\#Kleinanzeigen};
  \item links, e.g. \textit{http://t.co};
  \item smileys, e.g. \textit{:-P};
  \item slang, e.g. \textit{OMG}.
\end{enumerate}
according to the formal or lexical class which the tokens belonged
to. Additionally, we also considered as \textit{slang} spelling
variants of standard language words which could be regarded as their
colloquial equivalents. Such words included cases like, for example,
\textit{ned} instead of \textit{nicht} or \textit{grad} instead of
\textit{gerade} and were marked both as \textit{misspelling}s and
\textit{slang} in our classification. A detailed statistics on the
subgroups of class 3 is shown in Table \ref{table:style}:
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Stylistic specifics of text genre''\label{table:style}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    @-tokens & 13.02 & 20.23 & 16.19 & 21.91\\
    hashtags & 7.35 & 6.18 & 13.06 & 10.59\\
    links & 2.43 & 0.4 & 4.89 & 6.07\\
    smileys & 2 & 0.73 & 6.88 & 1.2\\
    slang & 16.22 & 5.27 & 6.94 & 4.77\\\hline
    {\bfseries total} & 41.02 & 32.81 & 47.96 & 44.54\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

A striking outlier of 16.22~\% for slang tokens in column 1 is
explained by the fact that the word ``RT'' which occurred 1,235 times
in our data and was by far the most frequent OOV in analyzed texts,
was recognized as out-of-vocabulary by \texttt{hunspell} but was not
deemed as such by \texttt{TreeTagger}. Luckily, such singleton cases
were rather rare exceptions during the analyses and did not affect
much the remaining classes of OOVs, so that distributions of unknown
tokens for both tools were more or less similar or at least comparable
to a certain extent.

But an even more important conclusion which can be drawn from the
analyzed data is the fact that for both \texttt{TreeTagger} and
\texttt{hunspell}, Twitter-specific phenomena like special tokens,
colloquial expressions or sloppily typed words accounted for more than
a half of all unknown words found in data. Since different classes of
these Twitter phenomena were formed by different processes and also
have different degrees of ambiguity, we will suggest different
normalization procedures for each of them in the next Section.

\section{Text Normalization Procedure}\label{normalization}

\subsection{Replacement of Twitter-Specific Phenomena}
Though hashtags are almost ubiquitous in today's tweets and one would
probably think that they are as old as Twitter itself, an admirable
truth would be that hashtags actually existed long before
Twitter-service was introduced. It were Internet Relay Chats (IRC) --
the forebears of all modern social networks -- which used the hash
sign to mark channel names. And it was Chris Messina, a renowned
advocate of open-source community, who borrowed this idea and brought
it to Twitter. The ``hash godfather'', as Chris Messina describes
himself, sent out a Twitter message suggesting that the pound sign
coul be used for organizing groups on Twitter in August 2007. But a
noticeable dissemination of this ``novelty'' among the Twitter
community only happened a few years later, in 2010. ``I would say 2010
was really the year of the hashtag.'' claims Ginger Wilcox, a founder
of the Social Media Marketing Institute (cf. Parker, \cite{parker}).

Luckily for us, both hashtags and @-names which are another
omnipresent phenomenon in Twitter have very clear formal criteria by
which these tokens could be identified and captured. The existence of
such formal cues suggests that these phenomena could best be handled
by means of rule-based methods. A most appropriate of such methods
seemed to be an analog of finite-state transducers (FST).

%
\subsection{Restoration of Umlauts}
%
\subsection{Squeezing of Elongated Words}
%
\subsection{Translation of Slang Idioms}
%
\section{Evaluation}\label{evaluation}

@@

\bibitem [2011]{parker}
  Parker, A.: %
  Twitter's Secret Handshake. %
  The New York Times, Page ST1, (June 10, 2011)
